---
title: "Reproduction"
format: html
editor: visual
---

## Dating a vegetarian?

Original paper by Adamczyk, D., Nezlek, J. B., & Maison, D. (2024)

Reproduction by Vijlbrief, A. J., & Zweistra, A. (2025)

Adamczyk, D., Nezlek, J. B., & Maison, D. (2024). Dating a vegetarian? Perception of masculinity, attractiveness, and the willingness to date vegetarians. *Social Psychological Bulletin*, *19*, 1-25.

Open-data: [https://osf.io/gek2d/](https://osf.io/gek2d/?view_only=)

### Set-up

Below is the code used to load the needed packages and read in the datafile.

```{r}
library(tidyverse)
library(haven)
library(lsr)
library(emmeans)

data <- read_sav("Dating a vegetarian_dataset.sav")
```

### Reproduction

The aim of this project is to reproduce target findings of the paper "Dating a vegetarian? Perception of masculinity, attractiveness, and the willingness to date vegetarians." by Adamczyk et al. (2024). In this reproduction we will limit ourselves to reproducing the series of ANOVA's that are the main interests in the original paper. These were a series of 4 (conditions) by 2 (sex) two-way ANOVA's testing effects on general attitudes, willingness to date, gender congruence, possession of masculine traits and the possession of feminine traits. Bonferroni corrected post-hoc comparisons were performed for all ANOVA's. In addition to this, tests of contrasts of the combined vegetarian conditions against the control condition were performed for all ANOVA's except the general attitudes test. A reason for this omission was not given. Adamczyk et al. used these results to support three out of their four main conclusions. The fourth conclusion was based on a series of ANCOVAs which we will not reproduce.

### **General data confirmation**

In this section we aim to verify general information about the sample as stated in the paper. According to the method section of the paper the sample comprises 404 heterosexual participants of which 204 are women and 200 are men. The age range of the study was 18-82 (M = 45.67, SD = 14.94). Additionally we examine the structure of their dietary variable.

```{r}
#verifying the number of participants (observations)
unique(data$Id)
#404 rows but from the participant id numbers it seems at least 509 observations were made

#verifying the number of women and men
#No codebook was provided next to the data. However, by examining the variables individually, labels for each of the categories were found in the tibble data.
data %>% count(sex) #204F, 200M 

#verifying heterosexual orientation of all participants
unique(data$orientation) #all heterosexual

#Examining labels of the diet variables (condition/COND)
#types of conditions
unique(data$condition)
#reports 5 conditions not four; control, VEGAN, eco, health and ethics
unique(data$COND)
#reports 4 conditions; control, eco, ethics, health

#verifying the age range and descriptives
data %>% count(age)
#categorical variable made up of different age ranges, ranging from 18-55+
```

From this primary exploration of the data we found most information provided in the paper to be accurate. The sample size and gender division were correct. We noted three points where the transparency of the paper could be improved. Firstly, from the respondent number (Id), we believe that at least 509 participants took part in this study. In their paper, Adamczyk et al. do not report any observations that were excluded from their analysis. We believe they excluded at least 105 participants due to unknown reasons and think the transparency of their research would be improved by reporting on this. Secondly, in their analyses, Adamczyk et al. use 4 conditions for the dietary variable (control, health, eco and ethics). However, in our exploration we found that there was originally a fifth condition (vegan). The researchers do not report on as to why this condition was not used in their analysis. Lastly, we were unable to verify the age range and its mean and standard deviation due to the structure of the variable. The variable was structured as categorical with different age groups. We were able to verify that the youngest participants were 18 but the maximum age, mean and standard deviation were not eligible for verification.

### Verification analysis

Adamczyk et al. performed their analysis through a series of ANOVAs and their post-hoc and contrast analyses. They do not report which statistical program they use to perform their analyses. We inferred that they used SPSS as the data file in their repository was a .sav file. We have no further information on which version of SPSS was used or what functions were utilized during the analysis.

The main analysis existed of five two-by-four ANOVAs to examine the main effects of diet and sex and their interaction effect. They did this for the following variables: ‘general likeability’ (termo), ‘willingness to date’ (dating_mean), ‘gender congruence’ (gender_cong), ‘masculine traits’ (masc_traits) and ‘feminine traits’ (fem_mean). They reported their ANOVAs through an F-test with p-value and an effect size by means of a partial eta squared. There was no information on the structure of their ANOVA tests, as such we decided to form the ANOVAs with the following formula: \~ factor(COND) \* factor(sex). We decided to have the variables function as factors and to multiply the dietary variable by the gender variable (instead of reversed). In theory, the order of the variables should not matter but through testing we found small differences in the main effects depending on which of the two variables was first in the formula. There was no clear pattern to indicate which of the two structures the researchers used in their analysis however, since they often reported the main effect of diet before that of sex, we decided on the aforementioned structure.

For their post-hoc test they report using the Bonferonni correction, no further elaboration is made about the method of post-hoc testing. As such, we were forced to infer how the post-hoc tests were performed. We decided to use pairwise t-tests with the Bonferroni correction as our form of post-hoc testing. In their results section they report whether there are any significant differences for any of the dietary conditions and the corresponding p-values. This is done for all variables except the general attitude 'termo' variable where the post-hoc test is reported per diet *per gender*. The reason for this incongruence is not reported. It could be that significant gender effects were only found for the termo variable but that is speculation. The transparency of the paper could have been improved by providing reasons for the alternative analysis/exclusion of the gender variable.

Adamczyk et al. provided weights for the contrasts in their results section leading us to use the 'emmeans' package for the contrast analyses. They reported contrasts for all analyzed variables expect for general attitudes. No reason was provided as to why they chose not to perform and/or include a contrast test for this variable. For completion's sake we chose to do the contrast test but are consequently not able to compare it to any result from the paper.

#### ANOVA for General attitudes

```{r}
#specifying the ANOVA design and requesting the results
anova_termo <- aov(termo ~ factor(data$COND) * factor(data$sex), data = data)
summary(anova_termo)

#requesting the effect size
etaSquared(anova_termo)

#post-hoc comparisons through pairwise t-tests for each dietary condition per sex
x <- vector(mode='list', length=4)
for (i in 1:4) {
  z <- data %>% filter(COND == i )
  x[[i]] <- pairwise.t.test(z$termo, z$sex, p.adj='bonferroni')
}
print(x)

#Contrasts
em <- emmeans(anova_termo, ~factor(COND))

(contrast_termo <- contrast(em, method = list(
  "-Control + Veg" = c(-3,1,1,1)
)))

```

The results of the ANOVA for the variable of General Likeability (termo) show only small differences in values. Similar to the original paper we found a significant interaction and diet effect and an insignificant effect of sex with small effect sizes. Post-host tests show significant differences between sexes for the ethics dietary condition. The test of contrasts shows a significant difference between the control condition and the combined vegetarian conditions, for unknown reason Adamczyk et al. did not perform a test of contrasts for this specific variable so no comparison can be made.

#### ANOVA for Willingness to date

```{r}
#specifying the ANOVA design and requesting the results
anova_will <- aov(dating_mean ~ factor(data$COND) * factor(data$sex), data = data)
summary(anova_will)

#requesting the effect size
etaSquared(anova_will)

#post-hoc comparisons through pairwise t-tests for each dietary condition
pairwise.t.test(data$dating_mean, data$COND, p.adj='bonferroni')

# Contrasts
em <- emmeans(anova_will, ~factor(COND))

(contrast_will <- contrast(em, method = list(
  "Control - Veg" = c(-3,1,1,1)
)))
```

The results of the ANOVA for the variable of Willingness to date (dating_mean) show mostly small differences in values with one big difference in p-value for the main effect of sex. Adamczyk et al. report a p-value of p = .383 whereas we find a p-value of p = .322. A difference of .061 is quite large however since it does not create a significance reversal it did not raise large concerns because the conclusions remain the same. It does however negatively impact the credibility of the published paper as no clear explanation could be found for the difference. Similar to the original paper we found a significant effect for diet only, with a small effect size. Post-host tests shows a significant difference between the control and the health condition and the test of contrasts shows a significant difference between the control condition and the combined vegetarian conditions.

#### ANOVA for Gender Congruence

```{r}
#specifying the ANOVA design and requesting the results
anova_gendCongr <- aov(gender_cong ~ factor(data$COND) * factor(data$sex), data = data)
summary(anova_gendCongr)

#requesting the effect size
etaSquared(anova_gendCongr)

#post-hoc comparisons through pairwise t-tests for each dietary condition
pairwise.t.test(data$gender_cong, data$COND, p.adj='bonferroni')

#Contrasts
em <- emmeans(anova_gendCongr, ~factor(COND))

(contrast_gendCongr <- contrast(em, method = list(
  "-Control + Veg" = c(-3,1,1,1)
)))
```

The results of the ANOVA for the variable of Gender congruence (gender_cong) show only small differences in values. Similar to the original paper we found a significant effect for diet and sex but no significant interaction effect. Post-host tests shows a significant difference between the control condition and the health/ethics conditions as well as a significant difference between the health and eco condition. The test of contrasts again shows a significant difference between the control condition and the combined vegetarian conditions.

#### ANOVA for Possession of masculine traits

```{r}
#specifying the ANOVA design and requesting the results
anova_masc <- aov(masc_traits~ factor(data$COND)*factor(data$sex), data = data)
summary(anova_masc)

#requesting the effect size
etaSquared(anova_masc)

#post-hoc comparisons through pairwise t-tests for each dietary condition
pairwise.t.test(data$masc_traits, data$COND, p.adj='bonferroni')

# Contrasts
em <- emmeans(anova_masc, ~factor(COND))

(contrast_masc <- contrast(em, method = list(
  "-Control + Veg" = c(-3,1,1,1)
))) 
```

The results of the ANOVA for the variable of Masculine traits (masc_traits) show important differences in values. The paper reported a significant effect of diet and sex with no interaction effect. However, our analysis indicated only diet to be significant and not sex. It is unclear how such a big difference in p-value can arise (Adamczyk et al. p = .014, Vijlbrief & Zweistra p = 0.064) and the difference does create a significance reversal. Post-host tests shows a significant difference between the control and the health condition and no further significant differences. The test of contrasts shows no significant differences either which is in line with the results of the publication.

#### ANOVA for Possession of feminine traits

```{r}
#specifying the ANOVA design and requesting the results
anova_fem <- aov(fem_mean ~ factor(data$COND)*factor(data$sex), data = data)
summary(anova_fem)

#requesting the effect size
etaSquared(anova_fem)

#post-hoc comparisons through pairwise t-tests for each dietary condition
pairwise.t.test(data$fem_mean, data$COND, p.adj='bonferroni')

# Contrasts
em <- emmeans(anova_fem, ~factor(COND))

(contrast_fem <- contrast(em, method = list( 
  "-Control + Veg" = c(-3,1,1,1)
)))
```

The results of the ANOVA for the variable of Feminine traits (fem_mean) show only small differences in values. Similar to the original paper we found a significant effect for diet and sex but no significant interaction effect. Post-host tests shows a significant difference between the control condition and the health condition as well as a significant difference between the health and eco condition. The test of contrasts shows no significant difference between the control condition and the combined vegetarian conditions which is in line with the results of the publication.

### Comparing the values 

All the previously mentioned ANOVA values are illustrated in the section below. We created histograms of the absolute differences between the F/P/Eta-values as reported in the published paper and ours.

```{r}
#Overview of differences

#F-values
  #storing the differences in F-values per variable as a dataframe 
Fvalues <- data.frame(
  diet = c(-0.01, 0.03, -0.19, +0.04, +0.05),
  sex = c(NA, -0.038, -0.09, -0.11, -0.21),
  interaction = c(0, 0, 0, 0, 0)
)

#creating a vector with all F-value differences
FCvalues <- c(-0.01, 0.03, -0.19, +0.04, +0.05, NA, -0.038, -0.09, -0.11, -0.21)

#creating a histogram of absolute differences in F-values
hist(abs(FCvalues), 
     col = 'lightblue',
     main = "Absolute differences in F-values",
     xlab = "Absolute difference")

#P-values
  #storing the differences in p-values per variable as a dataframe 
Pvalues <- data.frame(
  diet = c(0, -0.002, 0, -0.01, 0),
  sex = c(NA, -0.061, 0, +0.050, NA),
  interaction = c(0, 0, 0, 0, 0)
)

#creating a vector with all p-value differences
PCvalues <- c(0, -0.002, 0, -0.01, 0, NA, -0.061, 0, +0.050, NA)

#creating a histogram of absolute differences in p-values
hist(abs(PCvalues), 
     col = 'hotpink',
     main = "Absolute differences in p-values",
     xlab = "Absolute difference")

#Partial eta-squared values
#storing the differences in Eta-values per variable as a dataframe 
Etavalues <- data.frame(
  diet = c(0, 0, 0, 0, 0),
  sex = c(NA, -0.001, 0, 0, 0 ),
  interaction = c(0, 0, 0, 0, 0)
)

#vector with all eta-value differences
EtaCvalues <- c(0, 0, 0, 0, 0, NA, -0.001, 0, 0, 0, 0, 0, 0, 0, 0)

#creating a histogram of absolute differences in Eta-values
hist(abs(EtaCvalues), 
     col = 'palegreen2',
     main = "Absolute differences in Partial Eta-squared values",
     xlab = "Absolute difference")


```

From the histogram regarding F-values (blue) it can be seen that only small differences in F-values were observed (none higher than 0.25). These small differences could be due to differences in rounding in different coding programmes used for the analyses. Similar results can be observed for the Partial Eta-Squared values (green), where the largest observed difference was 0.001. The histogram for p-values (pink) showed a more interesting result. Two p-values had a difference of more than 0.04. These are of interest as they can more easily lead to different conclusions on significance and interpretations in the final conclusions of an analysis. As discussed, one of these p-values occurred in the variable Willingness to date but did not lead to differences in significance or conclusions. The other p-value occurred in the variable Masculine traits where it did lead us to a different conclusion. Adamczyk et al. concluded a significant effect of sex whereas we did not.

## Conclusion and evaluation

### Conclusions 

Overall most of the original papers findings were able to be reproduced. Most p-values and F statistics were off only by a marginal amount, which is probably due to rounding differences between the software used. However two p-values differed by .05 or more between our result and the published value. This is a serious error. For one of the results this means the value was interpreted as significant in the published paper while according to our calculations this is a non-significant result. It is unclear to us how this error made it into the final paper. It is possible that there was an error in the process of copying the value from the statistical software to the text editor as it only differs in one digit (0.064 to 0.014). The conclusions made by the original authors still largely hold up. In terms of significance only the conclusion about the effect on possession of masculine traits should be disregarded based on the p-value discrepancy. On the other analyses we get the same conclusions as the original authors.

Our goal was to verify the first three of the main conclusions from the publication: 1. Being a vegetarian made a person less attractive as a potential dating partner. 2. The control diet was evaluated most positively, and the vegetarian for health reasons the least positively (least dateable, least gender congruent, least masculine and feminine). 3. Men who were vegetarians for ethical reasons were perceived less positively than women who were vegetarians.

We were able to verify conclusions 1 and 3. Conclusion 1 was verified with significant a contrast analysis. Our results showed that willingness to date was lower for the combined vegetarian condition than the control condition (section: ANOVA for Willingness to date). Conclusion 3 was verified through a significant post-hoc test of differences between general likeability per diet and sex (section: ANOVA for General Likeability). We do believe that these conclusions might have benefitted from being worded slightly differently as we were unsure whether "perceived less positively" and "less attractive" referred to the variable of general likeability and/or willingness to date. We were unable to verify conclusion 2. We performed the post-hoc tests that were intended to support this conclusion (see section: Verification analysis) but did not reach the same results. The post-hoc tests did show that control condition had the highest mean in all variables except feminine traits. However, it also showed that control did not differ significantly from eco in any variable nor from ethics in any variable other than gender congruence. Similarly, health did indeed have the lowest mean in all mentioned variables but often only differed significantly from the control condition. Therefore, we do not feel that the results found through the described analysis warrant this conclusion.

### The process 

During the reproduction process we ran into several issues we'd like to address as considerations for making scientific work more reproducible. First of all, we want to thank the authors for making their data openly available. However, for their analyses to be truly reproducible we would also need more information on the statistical software used and the analyses themselves. The authors do not mention what program they used (although the .sav data file format implies they used IBM SPSS). Ideally, a syntax file (in the case of a program like SPSS, JASP or STATA) or a code file could be provided. This ensures strict reproducibility even though authors might be forced to leave out important information in the draft paper as it can be inferred from the code. Other than this, a full list of the software versions and package versions should also ideally be provided. This is somewhat less important, as these are unlikely to result in large differences, but is nevertheless essential to make sure the code remains usable in the long term. Furthermore, it is very important to include a complete codebook. A codebook was provided for this dataset on the OSF but it was not complete and lacked certain labels. Furthermore, information about how certain scale scores were constructed was absent. For example, in the dataset the variables "beh_1"to "beh_10" are present, alongside the general "beh" variable. It seems these variables were added up to make the scale score, but this can be an important methodological decision that should be explained in the text. Similarly, the reproducibility would benefit from specifying which version of the variables were used in the analysis process. The "dating_mean" variable that we used in our analysis is a mean variable of "beh", this use of mean data instead of summed data might cause small differences in outcomes. Lastly, we think that the reproducibility of this paper can be improved most easily through increased transparency. In our initial exploration of the data we noticed multiple mismatches between what was reported in the paper and what we observed from the data. It appeared as though 509 participants were observed instead of the reported 404 and as if there had initially been five dietary conditions (including vegan) instead of four. Reasons for these discrepancies may have been rather simple such as that the 105 unknown participants did not meet the sexuality or age requirements, however, since there was no reference to these participants in the published paper this is pure speculation. Furthermore, without any information on why the extra dietary category was not included it might lead reproducers to believe this was omitted because the effects were insignificant. This uncertainty and speculation can easily be prevented by briefly explaining why a decision was made. This issues continued throughout the paper as decisions such as why not to perform a contrast test for general likeability or why to perform a different post-hoc test were not elucidated. Similarly, further explanation on what type of post-hoc tests were used or the structure of the ANOVAs would have meant that we could have made fewer assumptions in our verification process which would mean that the reliability of the results would have increased. Therefore, the paper could have greatly benefitted from being more transparent about these decisions.

## References 

Adamczyk, D., Nezlek, J. B., & Maison, D. (2024). Dating a vegetarian? Perception of masculinity, attractiveness, and the willingness to date vegetarians. *Social Psychological Bulletin*, *19*, 1-25.

## AI statement 

No AI tools were used in creating this work. All text and code is original work by A. J. Vijlbrief and A. Zweistra.
